---
title: "is_valid"
author: "Team 6"
date: "10/31/2014"
output: html_document
---
#fuction 2: is_valid (Sophie)
## Validate the graph object to ensure that it meets all requirements - 
#Check that object is a list of lists. Check if there are names for the primary list that they are all unique. 
#Check that each secondary list contains only edges and weights vectors that are of the appropriate type. 
#Check that there are not any edges to non-existent vertices. 
#Check that all weights are not less than or equal to 0. 
#Check that every edge has a weight.

```{r}

library(testthat)
library(methods)

is_valid=function(input){

  # Testing if the input has a list object
if(typeof(input[[1]])=="list"){
  t1<-TRUE
}
  else {
  t1<-FALSE
  }

 # Testing if the names of the lists in object are all unique
names=names(input)
d_names=unique(names)
if(length(names)==length(d_names)){
  t2<-TRUE
}
else{
  t2<-FALSE
}

 # Testing if the object has both weights and edges as lists
t3_obj1<-grep("weights",names)
t3_obj2<-grep("edges",names)
if(sum(t3_obj1)>0 & sum(t3_obj2)>0){
  t3<-TRUE
}
else{
  t3<-FALSE
}

 # Testing if the contents in the lists have the mode of numeric
if(is.numeric(input[[1]]$edges)==TRUE & is.numeric(input[[1]]$weights)==TRUE){
   t4<-TRUE
 }
else{
  t4<-FALSE
}

# Testing that there are not any edges to non-existent vertices. 
n=length(input)
temp=NULL
t_obj1=list()
for(i in 1:n){
t_obj1[[i]]<-input[[i]]$edges
temp=c(temp,t_obj1[[i]])
}
t_obj2=(1:n)

if(order(unique(temp))==order(unique(t_obj2))) {
  t5<-TRUE
}
else{
  t5<-FALSE
}
# Testing if the weights are greater than 0.
if(input[[1]]$weights>0){
  t6<-TRUE
}
else{
  t6<-FALSE
}


# Testing if the object passes all tests above
each<-cbind(t1,t2,t3,t4,t5,t6)
if(sum(which(each==FALSE))>0) {
  print(FALSE)
}
else{
  print(TRUE)
}

}

```

### Testing

```{r}
basepath<-getwd()
newpath<-paste0(basepath,"/tests")
setwd(newpath)
test_file("test_is_valid")
testthat(is_valid)





```
