---
title: "is_valid"
author: "Team 6"
date: "10/31/2014"
output: html_document
---
#fuction 2: is_valid (Sophie)
## Validate the graph object to ensure that it meets all requirements - 
#Check that object is a list of lists. Check if there are names for the primary list that they are all unique. 
#Check that each secondary list contains only edges and weights vectors that are of the appropriate type. 
#Check that there are not any edges to non-existent vertices. 
#Check that all weights are not less than or equal to 0. 
#Check that every edge has a weight.

```{r}

library(testthat)
library(methods)


```

### Function

```{r}

is_valid=function(input){

# Testing if the input has a list object

if(as.character(typeof(input[[1]])=="list")=="TRUE"){
  t1<-TRUE
}
  else {
  t1<-FALSE
  }

# Testing if the names of the lists in object are all unique

if(as.character( length(names(input[[1]]))==length(unique(names(input[[1]]))) )=="TRUE"){
  t2<-TRUE
}
else{
  t2<-FALSE
}

 # Testing if the object has both weights and edges as lists

if(as.character( sum(grep("weights",names(input[[1]])))>0 & sum(grep("edges",names(input[[1]])))>0 )=="TRUE"){
  t3<-TRUE
}
else{
  t3<-FALSE
}

 # Testing if the contents in the lists have the mode of numeric

if(as.character(is.numeric(input[[1]]$edges)==TRUE & is.numeric(input[[1]]$weights)==TRUE)=="TRUE"){
   t4<-TRUE
 }
else{
  t4<-FALSE
}

# Testing that there are not any edges to non-existent vertices. 
if( is.numeric( length(input[[1]]$edges))==TRUE){
  
n=length(input[[1]]$edges)
  if(n>0){
    temp=NULL
    t_obj5_1=list()
    t_obj5_2=(1:n)
              for(i in 1:n){
              t_obj5_1[[i]]<-input[[i]]$edges
              temp=c(temp, t_obj5_1[[i]])
              } # end of for loop
    t_obj5= as.character( order(unique(temp))==order(unique(t_obj5_2)) )
        } #end of if condition
  else {
   t_obj5=as.character(TRUE) 
        }


  if(t_obj5=="TRUE") {
    t5<-TRUE
  }
  else{
    t5<-FALSE
  }


}
else{
  t5<-FALSE
}


# 6: Testing if the weights are greater than 0.
if( is.numeric( length(input[[1]]$edges))==TRUE){


    if( length(input[[1]]$weights)==0 ){
                                     t6<-TRUE
                                     }
    
    else{
            t_obj6<-as.character(input[[1]]$weights>0)
            if(table(t_obj6)[[1]]==table(rep("TRUE",length(input[[1]]$weights)))[[1]]){
                            t6<-TRUE
                            }
                        else{
                            t6<-FALSE
                            }
            
      }
  }

else{
  t6<-FALSE
    }

# Testing if the object passes all tests above
each<-cbind(t1,t2,t3,t4,t5,t6)
if(sum(which(each==FALSE))>0) {
  print(FALSE)
}
else{
  print(TRUE)
}
   print(each)  ## This should be removed before submission. For now, it shows the result of each test.
}



```



### Testing

```{r}
basepath<-getwd()
newpath<-paste0(basepath,"/tests")
setwd(newpath)
test_file("test_is_valid")
testthat(is_valid)

```

In case you are trying each object, use the codes below. Create g1:g6 and bad_g1:g3 first.
```{r}
is_valid(g1)
is_valid(g2)
is_valid(g3)
is_valid(g4)
is_valid(g5)
is_valid(g6)
is_valid(bad_g1)
is_valid(bad_g2)
is_valid(bad_g3)
```
