---
title: "hw2"
author: "Radhika Anand, Sophie Lee, Minjung Park, Kungang Zhang"
date: "10/4/2014"
output: html_document
---

```{r}
load(file="lq/lq_data.Rdata")
load(file="dennys/dennys_data.Rdata")
source("check_packages.R")
check_packages(c("httr","XML","jsonlite","stringr","fields"))
```

Task 1:

Data scraping for La Quinta is a 2 stage process and the code for the same is divided into 4 scripts. In the first stage we download some basic information about each La Quinta from a single page and in the second stage we access each La Quinta's webpage separately to get some specific information about each.
 
Stage 1 scripts: "lq_list_download.R" & "lq_list_parse.R"
Stage 2 scripts: "lq_hotel_download.R" & "lq_hotel_parse.R"
 
In the first script, "lq_list_download.R", we download and save the page, "www.lq.com/en/findandbook.by_interactive_map.html", locally, because this one single page has basic information about each La Quinta. In "lq_list_parse.R" we access this page from its local copy and parse the basic information for each La Quinta from the source code using regular expressions for string matching. In particular, we get "title", "innNumber", "latitude", "longitude", "imagePath", "isInnAndSuites", "street", "street2", "city", "stateProv", "postalCode" and "countryDisplay" and save them in a dataframe called "dataframe". We then drop "imagePath" and "isInnAndSuites".  We then drop "imagePath" and "isInnAndSuites" as we think they are not required. 
 
Further, in the script "lq_hotel_download.R", we access each La Quinta separately by its "InnNumber" using the following url, "http://www.lq.com/en/findandbook/hotel-details.[innNumber].address.html" where [innNumber] is replaced by the corresponding "innNumber" each time and save all the pages we get. 

Then in "lq_hotel_parse.R", we access each page locally and expand the existing "dataframe" to parse (using regular expressions for string matching) and include specific attributes such as "phone", "fax", number of "floors", "rooms" "suites", "check_in_time", "check_out_time" and access to specific amenities such as "spa", "internet", swimming "pool", "fitness" center. (Quotes are used to indicate the actual attribute names as used in the dataframe). We now have data for 880 La Quintas. We then remove the duplicates based on unique "innNumber" and are left with 875 unique ones. We further remove the data for Canada and Mexico as we need to test the claim only in US. We are then left with La Quintas. We name this final dataframe as lq.data.

Task 2: 

Sophie you can add your detailed write up here

Here, we are collecting the location data for Denny's. Since the location service provider for Denny's has a fairly high limit for radius, which is 3000 miles, we base the focal points for location collection in the following five states: Utah, DC, Kansas, Hawaii, and Alaska. 

Task 3: 

Testing Hedberg's claim that La Quinta is Spanish for "next to Denny's". We first calculate the distance between every pair of La Quinta and Denny's and store it in a matrix as shown below:
```{r}
#lq (lat,long) and denny's (lat,long), both stored in rows.
lqlatlong<-matrix(c(as.double(data.lq$longitude),as.double(data.lq$latitude)),ncol=2)
denlatlong<-matrix(c(as.double(as.matrix(dennys_data$longitude)),as.double(as.matrix(dennys_data$latitude))),ncol=2)

nlq=nrow(lqlatlong)
nden=nrow(denlatlong)

#a matrix of distance of all pairs of lq and denny's (nrow=lq, ncol=dennys)
distance<-NULL
distance<-rdist.earth(lqlatlong,denlatlong,miles=F,R=6371)
```
Function 'rdist.earth()' from package 'fields' is used to calculate distance on earth. Firstly, this function takes latitude and longitude values in degrees and not radians (we tested this out). Also, we return the distance in kms. Secondly, it assumes the earth’s radius to be 6378.388 km. According to Wikipedia this number seems to be the equatorial radius (the maximum radius). Because earth is not a perfect sphere, however, the radius declines as one moves to the poles reaching a polar minimum of about 6,357 km. The mean radius is 6371 km and thus we use R=6371 in our command. (inspiration for this claim: http://www.r-bloggers.com/great-circle-distance-calculations-in-r/)

We then plot a histogram of distance of each pair of LQ and Dennys and calculate mean distance as follows:

```{r,fig.align='center'}
#min and max distance
min=as.integer(min(distance))
max=as.integer(max(distance))

#hist and mean of distance matrix
hist(distance,breaks=100)
avedis<-mean(distance)
print(paste("The mean distance of (lq, denny's):",avedis))
```
We see that the histogram and mean give us summaries about all LQ Denny's pairs (no matter how far they are). We understand that this is not what we need to test Hedberg's claim. 

We then proceed to our next approach which accurately tests the claim. In this, we find the minimum distance of a Dennys from EVERY La Quinta (i.e. the distance between a La Quinta and its closest Denny's) and plot a histogram of these minimum distances. So if Hedberg's claim is true, there should be a denny's next to most LQs and the histogram should have high frequencies close to origin. We further define a variable for this measure (i.e. for the minimum distance of a Denny's from every La Quinta) and plot its Quantiles and CDF. From the plot of CDF, we can set any threshold distance for the term 'next to Dennys' and simply see the number/percentage of La Quintas that have a Dennys in that threshold.

```{r,fig.align='center'}
mindislq<-apply(distance,1,min)
#pdf("Min-distance",width=5.7,height=4.5)
hist(mindislq,breaks = seq(0,as.integer(max(mindislq))+1,0.1),freq = T
     ,xlim=c(0,20)
     ,xlab="Min-distance of (lq,denny's) for every lq (only show (0,20))(km)"
     ,ylab="frequency",main="Histogram of minimum distance of (lq,denny's) for every LQ")

#quantile function of the minimum distance of (lq,denny's) pair for every lq
qmindislq<-quantile(mindislq,probs = seq(0,1,0.01))

#pdf("quantile",width=5.7,height=4.5)
plot(seq(0,1,0.01),qmindislq,type="l",lwd = 2, col="red",
     xlab="Probability",ylab="Minimum distance(km)",main="Quantile curve")

#empirical cdf for the minimum distance
pmindislq<-ecdf(mindislq)
xp<-seq(0,max(mindislq)+1,length.out=1000)

#pdf("cdf",width=5.7,height=4.5)
plot(xp,pmindislq(xp),type="l",lwd = 2, col="green",
     ylab="Probability",xlab="Minimum distance(km)",main="CDF curve")
#dev.off()

#If we set the threshold to 0.5 km, the corresponding percentage and frequency:
print(paste("Per:",pmindislq(0.5)*100,"%",",","Freq:",pmindislq(0.5)*nlq))

#If increase the threshold to 1 km, the corresponding percentage and frequency:
print(paste("Per:",pmindislq(1)*100,"%",",","Freq:",pmindislq(1)*nlq))
```

We see that 20.07% of La Quintas have a Dennys within a 0.5 km radius from them and 29.87% of La Quinta's have a Dennys within a 1 km radius from them. At these percentages, the Hedberg's claim seems reasonable enough. However the claim is not always true.

We now plot the locations of La Quintas and Dennys for California and NC on their maps to pictorially understand the claim.

```{r,fig.align='center'}

suppressMessages(library(rgdal))
ogrInfo("/home/vis/cr173/Sta523/data/us-atlas/shp/us/","states-unfiltered")
states = readOGR("/home/vis/cr173/Sta523/data/us-atlas/shp/us/","states-unfiltered", stringsAsFactors=FALSE)

########### Save the location info of Denny's as Spatial Points
x=as.double(as.matrix(dennys_data$longitude))
y=as.double(as.matrix(dennys_data$latitude))
den_coord=cbind(x,y)
den = SpatialPoints(as.data.frame(den_coord)) 

########### Save the location info of La Quinta as matrix (Kungang already created this)
x=as.double(data.lq$longitude)
y=as.double(data.lq$latitude)
lq_coord=cbind(x,y)
lq = SpatialPoints(as.data.frame(lq_coord)) 

#########We pick California and North Carolina

###### California
dennys_ca = which(dennys_data$state=="CA")
den_coord_ca<-den_coord[dennys_ca,]
den_ca = SpatialPoints(data.frame(den_coord_ca)) 

laquinta_ca = which(data.lq$state=="CA")
lq_coord_ca<-lq_coord[laquinta_ca,]
lq_ca = SpatialPoints(data.frame(lq_coord_ca)) 

########## Now plot California
par(mar=c(2,2,1,1))
plot(states[states$STATE == "California",], col=c("lightgrey"), axes=TRUE)
points(den_ca, col="red", pch=16, cex=0.6)
points(lq_ca, col="blue", pch=16, cex=0.6)
par(mar=c(2,2,1,1))
legend("bottomleft",c("La Quinta","Denny's"),col=c("blue","red"),pch=16)


###### North Carolina
dennys_nc = which(dennys_data$state=="NC")
den_coord_nc<-den_coord[dennys_nc,]
den_nc = SpatialPoints(data.frame(den_coord_nc)) 

laquinta_nc = which(data.lq$state=="NC")
lq_coord_nc<-lq_coord[laquinta_nc,]
lq_nc = SpatialPoints(data.frame(lq_coord_nc)) 

########## Now plot North Carolina

plot(states[states$STATE == "North Carolina",], col=c("lightgrey"), axes=TRUE)
points(den_nc, col="red", pch=16, cex=0.6)
points(lq_nc, col="blue", pch=16, cex=0.6)
legend("bottomleft",c("La Quinta","Denny's"),col=c("blue","red"),pch=16)

```
From the maps we see that a lot of LQs have a Dennys close to them!!
