---
title: "hw3"
author: "Radhika Anand, Sophie Lee, Minjung Park, Kungang Zhang"
date: "10/19/2014"
output: html_document
---
## Task 1 : Geocoding

1) Load the biggish NYC parking violation datafile by using a (fread) function which is much faster and more convenient than a (read.table) function 

2) Extract the necessary variables

Using 'setnames' we changed some variable names. Then using the library 'dplyr', first we chose a part of [Violation.Precinct] data which is less than 34 by filtering. Second, by using a (mutate) function, we added two variables [House.number, Street.Name], erasing whitespace in the data of two variables. Then we took the non-empty data and chose [House.Number] data which has a "[0-9]+" pattern by using a (filter) function. Third, we combined the [Violation.Precinct] variable with the [House.Number,Street.Name] variable which is called "addr" by using transmute function and then lastly converted the case to lower.

3) Load 'pluto data' to geocode by using (readShapeSpatial) which reads data from a spatial dataframe to a shapefile.

We made a dataset called 'tax' which combines coordinates of 'pluto' data with address of it. Then we called the address of 'pluto' data 'addr' which is the same variable as we used in the dataframe above so as to implement inner_join on them.

3) Clearing messy data

First, we made a combined vector called 'namemap' which includes the full name ("east","west","street","avenue","av","avee","road","drive","terrace","bway","terrace"," terr","bway","1st","2nd","3rd","th ") and the short name ("e","w","st","ave","ave","ave","rd","dr","ter","brdway","ter"," ter","brdway","1","2","3"," ") and then, in the for loop, we replaced the full name with the short name in both 'tax' data from pluto and 'addr' data from park. We used 'str_match_all' to do so. So after cleaning a few words and removing suffix from numbers, we could atch upto 1.5 mil rows. 

```{r}
#setwd("~/Team6/HW3")

#Load necessary packages
source("check_packages.R")
check_packages(c("data.table","rgeos","ggmap","rgdal","maptools", "dplyr","stringr","lubridate"))

#Load the Biggish NYC parking violations datafile
park = fread("/home/vis/cr173/Sta523/data/parking/NYParkingViolations.csv",sep=",")

#Rename required columns
setnames(park, "Violation Precinct", "Violation.Precinct")
setnames(park, "House Number", "House.Number")
setnames(park, "Street Name", "Street.Name")


#Extract the necessary variables
addr = filter(park, Violation.Precinct <= 34) %>%
  mutate(House.Number = str_trim(House.Number), Street.Name = str_trim(Street.Name)) %>%
  filter(House.Number != "" & Street.Name != "") %>%
  filter(str_detect(House.Number,"[0-9]+")) %>%
  transmute(Violation.Precinct = Violation.Precinct, addr = paste(House.Number, Street.Name)) %>%
  mutate(addr = tolower(addr))

#rm(park)


#Load 'pluto' data to geocode
basepath<-getwd()
setwd("/home/vis/cr173/Sta523/data/parking/pluto/Manhattan/")
pluto<-readShapeSpatial("MNMapPLUTO")
setwd(basepath)
rm(basepath)

tax = cbind(data.frame(coordinates(pluto)), tolower(as.character(pluto$Address)))

names(tax)[3] = "addr"


#from full name to simplified version, becasue this make str_replace_all much easier
namemap<-rbind(c("east","west","street","avenue","av","avee","road","drive","terrace"," terr","bway","1st","2nd","3rd","th "), c("e","w","st","ave","ave","ave","rd","dr","ter"," ter","brdway","1","2","3"," "))
lnm<-ncol(namemap)

for (i in 1:lnm)
{
  tax$addr<-str_replace_all(tax$addr, namemap[1,i], namemap[2,i])
  addr$addr<-str_replace_all(addr$addr, namemap[1,i], namemap[2,i])
}

#Around 1.5 mil matches after this clean up
```

4) Next, we used a more comprehensive approach to clean the data where we saw, each time, the number of unmatched rows in pluto and then tried to clean the words in those rows. We did this but it was taking a lot of time and space and was increasing the matching by a very small and insignificant amount. Hence, we have commented this code for now. It is given below:

```{r}
#Better approach to clean up, but doesnt increase data points by a significant number. 
#We can do it but its takes a lot of time and space, so we have commented it for now.

#some house number has 0 at the beginning, the total number is about 1100

#hnvset<-str_detect(park$House.Number, "^0+([[:alnum:]]+)")
#hnpset<-str_detect(tax$addr, "^0+([[:alnum:]]+)")

#addrset<-str_detect(addr$addr, "^0+([[:alnum:]]+)")
#addrmat<-str_match(addr$addr[addrset], "^0+([[:alnum:]]+)")
#nset<-length(addrset)
#seqn<-c(1:nset)
#seqnmark<-seqn[addrset]#record the seq number of element detecting the pattern with first 0
#nseqnmark<-length(seqnmark)#nseqmark is 1109

#for(i in 1:nseqnmark)
#{
#  addr$addr[seqnmark[i]]<-str_replace(addr$addr[seqnmark[i]], addrmat[i,1], addrmat[i,2])
#}
#addrset1<-str_detect(addr$addr, "^0+([[:alnum:]]+)")

#However this part only increase ~100 data points in the final z

#after cleaning 1st, 2nd, 3rd, and ~th, the matching goes up to 1569809
#after delete first 0 in house number, the matching goes up to 1569912

#unique addr$addr
#univaddr<-unique(addr$addr)
#length(addr$addr)
#length(univaddr)
#unique z$addr
#unizaddr<-unique(z$addr)
#length(unizaddr)

#not.match<-anti_join(addr$addr,z$addr)
#don't work: Error in UseMethod("anti_join") : 
#no applicable method for 'anti_join' applied to an object of class "character"

#not.mav<-anti_join(addr,tax)
#not.map<-anti_join(tax,addr) #The addr only appear in tax
#taxdt<-as.data.table(tax) #need to convert tax into data.table, otherwise would report error
#not.mav<-anti_join(addr,taxdt) #The addr only appear in addr

#uni.not.mav<-unique(not.mav)
#uni.not.map<-unique(not.map)

#By inspecting data in not.map and not.mav, using format below, we can find how many data can be changed with each modification
#If the number to increase the valid data is small, then it is not worthy to loop over all of data once
#test<-str_detect(addr$addr, " terr")
#test[is.na(test)]<-F
#sum(test) #249

#test<-str_detect(addr$addr, " square$")
#test[is.na(test)]<-F
#sum(test) #821

#test<-str_detect(tax$addr, " square")
#test[is.na(test)]<-F
#sum(test) #92

#test<-str_detect(addr$addr, " sq$")
#test[is.na(test)]<-F
#sum(test) #4940

#test<-str_detect(tax$addr, " sq")
#test[is.na(test)]<-F
#sum(test) #93
```

5) Joining addr and tax using 'inner_join' and removing park and pluto

```{r}
# Join the dataframe
z = inner_join(tax,addr)

#plot(z$X1,z$X2,col=z$Violation.Precinct)

rm(park)
rm(pluto)
```

6) Plotting the precincts on map in different colors to get an idea of the boundaries. We see that the concentrated points are very good but we have a lot of outliers to get rid off.

```{r,fig.align='center'}
latlon=data.frame(cbind(z$X1, z$X2))
names(latlon$X1)="lon"
names(latlon$X2)="lat"
coord=SpatialPoints(latlon)
plot(coord, col=z$Violation.Precinct, pch=18, cex=0.5, axes=TRUE)
dim(latlon)
```

## Task 2 : Recreating NYC’s Police Precincts

1) We use the actual precinct numbers from NYPD website to subset the actual Manhattan precincts between o to 34. We got the list as 1,5,6,7,9,10,13,14,17,18,19,20,22,23,24,25,26,28,30,32,33,34 where Midtown South is 14, Manhattan North is 18, Central park is 22.

2) We got rid of outliers by taking the 90% quantile of lat/long pairs. Since points are numeric we can use the quantile function which sorts and removes 10% outliers for each precinct.

3) We then made separate convex hulls for each precinct (restoring the precinct number for later use) using the function 'gConvexHull' from 'rgeos' library. We then merged all the convex hulls and made a single geojson file called 'precinct.json' (using the function 'writeOGR') as below:

```{r}
object=SpatialPointsDataFrame(coords=latlon,data=data.frame(z$Violation.Precinct))

# obj is a list of all convex hulls. ch[[1]] has convex hull for precinct number 0, ch[[2]] for precinct 1 till ch[[35]] for precinct 34.

obj = list()
ch = list()
chl = list()
true_p=c(1,5,6,7,9,10,13,14,17,18,19,20,22,23,24,25,26,28,30,32,33,34)
j = 1

for(precinct in true_p)
{
  obj[j]=SpatialPointsDataFrame(coords=coordinates(object[object$z.Violation.Precinct==precinct,]),data=data.frame(object[object$z.Violation.Precinct==precinct,]$z.Violation.Precinct))
  
  p<-as.data.frame(object[which(object$z.Violation.Precinct==precinct),])
  z=data.frame(coordinates(obj[j]), p)
  q_lon<-matrix(quantile(z$X1, probs=c(0.05, 0.95), na.rm=TRUE))
  q_lat<-matrix(quantile(z$X2, probs=c(0.05, 0.95), na.rm=TRUE))
  pres<-which( q_lon[1,1]< z$X1 & z$X1<q_lon[2,1] & q_lat[1,1]< z$X2 & z$X2<q_lat[2,1])
  z<-z[pres,] 
  obj[j]=SpatialPointsDataFrame(coords=cbind(z$X1,z$X2), data=data.frame(z$z.Violation.Precinct)) 

  ch[j] = gConvexHull(obj[[j]])
  chl[j] = SpatialPolygonsDataFrame(ch[[j]],data=data.frame(precinct)) 
  j = j+1
}

# merge
merged = chl[[1]]

for(precinct in 2:length(true_p)) {
  #print(i)
  slot(slot(chl[[precinct]],"polygons")[[1]],"ID") = as.character(precinct)
  merged = rbind(merged,chl[[precinct]])
}



writeOGR(merged, "./out", "", driver="GeoJSON") # Creates out file, current version of 
# GDAL does not allow . in file names 
# so we have to rename the file afterwards
file.rename("./out", "./precinct.json")

```

To get rid of outliers in a better way, we plan to do the following (in the coming week): 

1) Check say 5 nearest neighbors of each point and if they all lie very far away then remove them but that is very complex

2) Remove 80% of the observations randomly (assuming 80% removes most outliers) and just use the left out 20% to make boundary and do this a 1000 times and take the intersection of all the 1000 polygons.
