---
title: "hw5"
author: "Team 6 - Radhika Anand, Sophie Lee, Minjung Park, Kungang Zhang"
date: "11/10/2014"
output: html_document
---


```{r}
#Load necessary packages first!
source("check_packages.R")
check_packages(c("parallel","xtable", "doMC", "foreach", "truncnorm", "iterators")) 
```

#IMPORTANT NOTE:
The .Rdata files are present in the hw5 folder so we can directly run the .RmD. But to reproduce the data, please run the 4 .R files before running the .RmD. The 4 .R files are: reject.R, mh_final.R, slice.R and R.R. 

In this assignment, we test the time difference between the single core and the multicore implementations of codes in R, required for running four samplers. 

# Samplers 

The four samplers constructed in testing the running time for single core and multi core are as follows. 

###NOTE:
We have used a general strategy common to all samplers that when mc=FALSE we use a single core and when mc=TRUE we use 2 cores if sample size is 10000 or less and 8 cores if sample size is greater than 10000.

## 1. Rejection sampling

Strategy:
1. We first sample uniformly from the X axis -- this is the x value of the envelop density. Since all the given test distribution functions have specified ranges, the X values for the envelop function is (min(range), max(range)). With the X values, we create an envelope function which is greater than or equal to the target distribution we wish to sample from at all points. 

2. We use a for loop function for each sampled value's rejection/acceptance. For each value from the target distribution, evaluated at the drawn X, is compared to that of the envelop function. Specifically, we calculate the ratio of each Y value drawn from the target distribution to the envelop function, which is the unifrom density. 

3. If the drawn Y value from the envelop function falls under the target distribution, then accept the value. Otherwise, throw away the Y value. This process is repeated for n times.

```{r}
reject = function(n, dfunc, range, mc){ 
  stopifnot(is.function(dfunc))
  stopifnot(is.numeric(n))
  stopifnot(is.logical(mc))
  stopifnot(is.vector(range))

  if(mc==FALSE){
    mc=1
  } else{
    if(n>10000)
    {
      mc=8
    } else{
      mc=2
    }
  }
  
  ## 1) generate values of Xs and Ys from uniform
 
  rejection=function(n,dfunc,range,mc){  
    proposal_x=runif(n, min(range), max(range))    
  
    ## 2) reject if Y is larger than the value from dfunc
    result_x=rep(0,n)
    result_x[1]=proposal_x[1]
    for(i in 2:n) {
      x=proposal_x[i]      
      y=dfunc(x) #target dist. evaluated at x
      ratio=y/1 # calculate the ratio  (density of unif. evaluated at x =1)
      if(runif(1,0,1)<ratio){ 
        result_x[i]=x
      } else{result_x[i]=result_x[i-1]} # end of acceptance condition
    } #end of rejection for loop
    return(result_x)
  }
  
  vec=unlist(mclapply(1:mc, function(x) rejection(ceiling(n/mc),dfunc,range,mc),mc.cores=mc))
  vec=vec[1:n]
  return(vec)
  
}  # end of function: reject

```

###Score table for rejection sampler when n = 1,000,000
The code to generate the following table is included in the reject.R file, which is saved under hw5 directory. 

```{r}
load("score_reject_data.Rdata")
options(warn=-1)
x=xtable(score_reject_data, type="html", digits=3)
options(warn=0)
```
```{r table1,results='asis'}
print.xtable(x,type="html")
```

### Singlecore timings (per iteration)- Rejection sampler
The code to generate the following table is included in the reject.R file, which is saved under hw5 directory. 

```{r}
load("reject_data.Rdata")
tt<-droplevels(reject_data[24,4])
tt<-as.numeric(levels(tt))

time_per_iteration=c(as.numeric(levels(reject_data[,4])), tt)

s=c("100", 
    "10000",
    "1000000",
    "10000000")
iteration=rep(s,6)
table=data.frame("distribution_names"=unlist(reject_data[,1]),"N_iteration"=iteration,
                time_per_iteration)
```
```{r table,results='asis'}
options(warn=-1)
x=xtable(table, type="html", digits=6)
options(warn=0)
print.xtable(x,type="html")

```
### Multicore timings (per iteration)- Rejection sampler

```{r}
load("reject_data_T.Rdata")
tt<-droplevels(reject_data_T[24,4])
tt<-as.numeric(levels(tt))

time_per_iteration=c(as.numeric(levels(reject_data_T[,4])), tt)

s=c("100", 
    "10000",
    "1000000",
    "10000000")
iteration=rep(s,6)
table=data.frame("distribution_names"=unlist(reject_data_T[,1]),"N_iteration"=iteration,
                time_per_iteration)

```
```{r table,results='asis'}
options(warn=-1)
x=xtable(table, type="html", digits=6)
options(warn=0)
print.xtable(x,type="html")
```

## 2. Metropolis Hastings sampling

Approach:
1) We have used metropolis random walk method with a normal proposal density.
2) We use the first 200 samples to tune the value of sigma square used in the normal proposal density. The tuning is such that the acceptance probability is between 0.3 to 0.4 percent. If acceptance probability is small, the sigma square is made smaller and vice versa. We have written a function called 'tuning' for this purpose. It returns the tuned value of sigma square.
3) We then implement the metropolis random walk in the function called sampling. Here, the standard metropolis random walk is implemented. That is if the sample from unif(0,1) is less than the acc prob then we accept the proposal, else reject it and keep the old sample.

The code is as below:

```{r}
library(xtable)

mh = function(n, dfunc, range, mc=FALSE){
  library(parallel)
  library(doMC)
  library(foreach)
  
  stopifnot(is.function(dfunc))
  stopifnot(is.numeric(n))
  stopifnot(is.logical(mc))
  stopifnot(is.vector(range))
  
  if(mc==FALSE){
    mc=1
  } else{
    if(n>10000)
    {
      mc=8
    } else{
      mc=2
    }
  }
  
  min=min(range)
  max=max(range)
  
  tuning=function(dfunc,min,max)
  {
    v=1.5
    tune=0.1
    while(tune<0.3 || tune>0.4){
      if(tune<0.1)
      {
        v=v/1.2
      } else 
      {
        if(tune>=0.1 && tune<0.3)
        {
          v=v/1.1
        } else 
          {
            if(tune>0.55)
            {
              v=v*1.25
            } else 
              {
                if(tune<=0.55 && tune>0.4)
                {
                  v=v*1.15
                }
              }
          }
      }
      
      count=0
      tunmat=matrix(0,200,1)
      tunmat[1]=runif(1,min,max)
      
      for(i in 2:200) 
      {
        prop=rnorm(1,tunmat[i-1],v)
        
        if(prop<min || prop>max)
        {
          tunmat[i]=tunmat[i-1] 
        } else{
          acprob=min(1,dfunc(prop)/dfunc(tunmat[i-1]))
          sam=runif(1,0,1)
          if(sam<=acprob)
          {
            tunmat[i]=prop
            count=count+1
          } else
          {tunmat[i]=tunmat[i-1]
          }
        }
      }  
      tune=count/200
    } 
    return(v)
  }
  v=tuning(dfunc,min,max)
  
  sampling=function(n,dfunc,min,max,v){
    
    x=matrix(0,n,1)
    x[1]=runif(1,min,max)
    
    for(i in 2:n) 
    {
      prop=rnorm(1,x[i-1],v)
      
      if(prop<min || prop>max)
      {
        x[i]=x[i-1]
      } else{
        accprob=min(1,dfunc(prop)/dfunc(x[i-1]))
        sam=runif(1,0,1)
        if(sam<=accprob)
        {
          x[i]=prop
        } else
        {x[i]=x[i-1]}
      }
    }
    return(x)}
  vec=unlist(mclapply(1:mc, function(x) sampling(ceiling(n/mc),dfunc,min,max,v),mc.cores=mc))
  vec=vec[1:n]
  return(vec)
}
```

### Table with time/iteration for single core and multicore and scores (SC and MC) for 1 mil samples.
Code is available in mh_final.R file in hw5 folder.

```{r}
load("MHtimes.Rdata")
options(warn=-1)
x=xtable(mainmh, type="html", digits=3)
options(warn=0)
```
```{r table44,results='asis'}
print.xtable(x,type="html")
```

## 3. Slice sampling

#Algorithm
The paper ``Radford Neal, the annals of stat., 03-slice sampling" shows a way to sample under the area of pdf function. The algorithm is as follows.

1) Starting from the support regime of the density function $df(x)$, outside which the value of the function is only zero, uniformly draw one sample $x_0$ from that regime.
2) Find the $df(x_0)$ and uniformly draw one sample from [$0$,$df(x_0)$] giving $y_0$.
3) Draw a horizontal line, and find the smallest regime where $df(x)>y_0$. And uniformly draw on sample from that regime giving $x_1$.
4) Draw a sample from vertical line segment [$0$,$df(x_1)$] giving $y_1$.
5) Repeat step 3) and 4) until the sample size of $\{x_i\}$ reaches given value.

##Realization
In step 3), finding a horizontal line segment with two end points giving $df(x_{left,i})=df(x_{right,i})=y_i$ is not easy. We need set a step length $w$. Starting from $x_i$, expand the regime in that step length in both direction while checking if the end points have reached outside of density curve. 

There is one problem in the fixed step length. If there exists a infinitly sharp mode, meaning the width of the tip goes to zero, $w$ may not well capture the regime at very narrow place. Then, if $x_i$ falls outside of that regime, we need to reject it and re-draw. The slice sampling can get stuck at this rejection sampling for extremely long time. To solve this problem, a threshold number for iteration in rejection sampling is set. 


```{r}

slice = function(n, dfunc, range, mc){
  stopifnot(is.function(dfunc))
  stopifnot(is.numeric(n))
  stopifnot(is.logical(mc))
  stopifnot(is.vector(range))
  
  if(mc==FALSE){
    mc=1
  } else{
    if(n>10000)
    {
      mc=8
    } else{
      mc=2
    }
  }
  
  check<-function(x)
  {
    if (is.finite(x) & is.numeric(x)) return(0)
    else
    {
      if (is.na(x)==TRUE) return(1)
      else
      {
        if (is.nan(x)==TRUE) return(2)
        else 
        {
          if ((is.infinite(x)==TRUE)) return(3)
          else return(4)
        }
      }
    }
  }
  
  sampling<-function(nc,dfunc,range,cores) 
  {
    rangec=range
    res<-matrix(NA,n,ncol=1,nrow=nc)
    w=abs(rangec[1]-rangec[2])/10
    y=0
    for (i in 1:nc)
    {
      res[i]=runif(1,rangec[1],rangec[2])
#       stopifnot(check(res[i])==0)
#       stopifnot(check(dfunc(res[i]))==0)
      k=0
      kk=0
      while (dfunc(res[i])<y) 
      {
        if (res[i]<res[i-1])
        {
          rangec[1]=res[i]
        }
        else
        {
          if (res[i]>res[i-1])
          {
            rangec[2]=res[i]
          }
        }
        res[i]=runif(1,rangec[1],rangec[2])
#         stopifnot(check(res[i])==0)
#         stopifnot(check(dfunc(res[i]))==0)
        k=k+1
        if(k>10)
        {
          kk=kk+1
         # print(c(k,kk,res[i],cores))
          rangec=range
          res[i]=runif(1,rangec[1],rangec[2])
#           stopifnot(check(res[i])==0)
#           stopifnot(check(dfunc(res[i]))==0)
          break
        }
      }
      y=runif(1,0,dfunc(res[i]))
      rangec=rep(res[i],2)
      while (dfunc(rangec[1])>y)
      {
        rangec[1]=rangec[1]-w
      }
      while (dfunc(rangec[2])>y)
      {
        rangec[2]=rangec[2]+w
      }
    }
    return(res)
  }
  
  cores=mc
  nc=ceiling(n/cores)
  sample<-unlist(mclapply(1:cores, function(z) sampling(nc,dfunc,range,cores),mc.cores = cores))
  sample=sample[1:n]
  return(sample)
}

```

### Score table for slice sampler when n = 1,000,000
The code to generate the following table is included in the slice.R file, which is saved under hw5 directory. 

```{r}
load("score_slice_data.Rdata")
options(warn=-1)
x=xtable(score_slice_data, type="html", digits=3)
options(warn=0)
```
```{r table4,results='asis'}
print.xtable(x,type="html")
```

###Singlecore (time/iteration) - Slice sampler
The code to generate the following table is included in the slice.R file, which is saved under hw5 directory. 
```{r}
load("slice_data.Rdata")
options(warn=-1)
x=xtable(slice_data, type="html", digits=3)
options(warn=0)
```
```{r table5,results='asis'}
print.xtable(x,type="html")
```
###Multicore (time/iteration) - Slice sampler
```{r}
load("slice_data_T.Rdata")
options(warn=-1)
x=xtable(slice_data_T, type="html", digits=3)
options(warn=0)
```
```{r table6,results='asis'}
print.xtable(x,type="html")
```




## 4. R sampling

Here we have used R's inbuilt sampling functions such as rbeta, rexp, rnorm, rtruncnorm etc. to sample from the given distributions. To sample from the truncated distributions we have used a slightly larger value of n and then returned only the samples that were in the range. We did not loop over the n to get exact number of samples as that was very time consuming. For mixture models, we sampled in the same percentage as was done in ethe function implementation. 

The code is present in the R.R file which generates the .Rdata file for the table. The code for R sampler is as follows:


```{r}
### R function 
library(xtable)
source("check_packages.R")
check_packages(c("parallel","xtable", "doMC", "foreach", "truncnorm", "iterators")) 

R = function(n,dfunc,range,mc){
  
  library(parallel)
  library(doMC)
  library(foreach)
  
  stopifnot(is.function(dfunc))
  stopifnot(is.numeric(n))
  stopifnot(is.logical(mc))
  stopifnot(is.vector(range))
  
  if(mc==FALSE){
    mc=1
  } else{
    if(n>10000)
    {
      mc=8
    } else{
      mc=2
    }
  }
  
  if(as.character(substitute(dfunc))=="dbetann"){
    return(unlist(mclapply(1:mc,function(x) rbeta(ceiling(n/mc),0.9,0.9),
                           mc.cores=mc)))}
  
  if(as.character(substitute(dfunc))=="dtnorm"){
    return(unlist(mclapply(1:mc,function(x) rtruncnorm(ceiling(n/mc),-3,3),
                           mc.cores=mc)))}
  
  if(as.character(substitute(dfunc))=="dtexp"){
    sample=unlist(mclapply(1:mc, function(x) rexp(ceiling((n/mc)*1.02),rate=1/3),
                   mc.cores=mc))
    return(sample[sample<6])
  }
  
  if(as.character(substitute(dfunc))=="dunif_mix"){
    return(unlist(mclapply(1:mc, function(x) c(runif(ceiling(0.6*n/mc),-3,-1),
                                                  runif(ceiling(0.1*n/mc),-1,1),
                                                  runif(ceiling(0.3*n/mc), 1,4)),
                           mc.cores=mc)))}
  
  if(as.character(substitute(dfunc))=="dtnorm_mix1"){
    return(unlist(mclapply(1:mc, function(x) c(rtruncnorm(ceiling(0.5*n/mc),0,10,2,2),
                                                  rtruncnorm(ceiling(0.5*n/mc),0,10,6,1)),
                           mc.cores=mc)))}
  
  if(as.character(substitute(dfunc))=="dtnorm_mix2"){
    return(unlist(mclapply(1:mc,
                           function(x) c(rtruncnorm(ceiling(0.45*n/mc),-4,4,-4),
                                         rtruncnorm(ceiling(0.45*n/mc),-4,4,4),
                                         rtruncnorm(ceiling(0.1*n/mc),-4,4,0,0.5)),
                           mc.cores=mc)))
  }
}

```

### Table with time/iteration for single core and multicore and scores (SC and MC) for 1 mil samples.
Code is available in R.R file in hw5 folder.

```{r}
load("Rtimes.Rdata")
options(warn=-1)
x=xtable(mainr, type="html", digits=3)
options(warn=0)
```
```{r table54,results='asis'}
print.xtable(x,type="html")
```


